On the marginal value of swimming in woodland caribou

### 2021-04-12 paper/manuscript.Rmd
new P1

The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to
approximately 300 woodland caribou. During routine fieldwork on May 30, 2017, we
observed an unmarked adult male caribou swim between two small islands (Figure 1
& 2), a distance of at least 470 m which took ~9 minutes (~52 m per minute).
Residents of Fogo Island confirmed this phenomenon and have also reported
regularly observing caribou swimming between islands. This observation prompted
us to investigate whether swimming behaviour was common among GPS radio-collared
caribou in the population [n = 29, for collaring details see, @Peignier2019]. We
identified an individual's swimming events as two consecutive GPS locations
(2-hour relocation rates) on different islands. Newfoundland typically
experiences pack ice in late winter and caribou ,may travel between islands by
walking over the ice. To identify swimming events, we restricted our dataset
between April 1 and December 31 and identified `r nevents` swimming events over
three years (Figure 2). Given the prevalence of swimming in this population, we
discuss the causes and consequences of this behaviour and propose future areas
of research.

In total, 12 of 29 collared female caribou swam between islands in the Fogo
Island archipelago. Of the caribou that swam, caribou swam on average
approximately `r neventidMean` (range = `r neventidMin`-`r neventidMax`) times
per year and remained on each island for approximately `r dur` (range = 
`r durMin`-`r durMax`) days before swimming again. Based on GPS data to infer
swimming, we separated the population into three apparent groups of individuals:
those that swam regularly between islands (n = 3, every ~`r indMean` days);
those that rarely engaged in swimming events (n = 9, every ~ `r ind2Mean` days);
and those that never swam (n = 17).

>>>>

Movement is costly. Even for adept swimmers like caribou (Rangifer tarandus), which have hollow hair shafts that aid in flotation, the energetic expenditure of swimming for terrestrial mammals is markedly higher than walking or running (Miller and Gunn 1985). The drivers that promote the decision to swim for terrestrial mammals, like caribou, are often related to the profitability of forage (Miller et al. 1977) and the risks of predation (Jeffery et al. 2007, Jordan et al. 2010). The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to approximately 300 woodland caribou, descendants of 26 individuals introduced from the Island of Newfoundland to Fogo Island between 1964–67 (Bergerud and Mercer 1989). During fieldwork on May 30, 2017, we observed an unmarked adult male caribou swim between two smaller islands (Figure 1) a distance of 470 m which took approximately 9 minutes. Given that swimming is energetically expensive, we ask how often do caribou swim between islands, how long do they occupy islands, and ultimately what motivates swimming? Here, we frame swimming and island residency as a model of patch use as described by the marginal value theorem (Charnov 1976) within the context of the forage limitation hypothesis. For caribou, the forage limitation hypothesis is the constraints on food availability associated with changes in abundance through time (Klein 1991, Schaefer et al. 2016). Here, we integrate our understanding of forage limitation in caribou populations by using island residency as a proxy for time in patch and island size as a proxy for energy intake in a patch (Figure 2).

`r nevents` 
In total, 12 of 29 collared female caribou swam between islands in the Fogo
Island archipelago. Of the caribou that swam, caribou swam on average
approximately `r neventidMean` (range = `r neventidMin`-`r neventidMax`) times
per year and remained on each island for approximately `r dur` (range = 
`r durMin`-`r durMax`) days before swimming again. Based on GPS data to infer
swimming, we separated the population into three apparent groups of individuals:
those that swam regularly between islands (n = 3, every ~`r indMean` days);
those that rarely engaged in swimming events (n = 9, every ~ `r ind2Mean` days);
and those that never swam (n = 17).
